Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-09-12 paper/broom.tex
Contributions added.
Sec 2 full content.

\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
\usepackage[breaklinks]{hyperref}

%Formatting Commands
%-------------------
\newcommand{\C}[1]{\code{#1}}
\renewcommand{\k}[1]{\mathtt{#1}} % Code font in math mode.
\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\conj}{\wedge}
\newcommand{\disj}{\vee}
\newcommand{\seqn}[1]{\overline {#1}}
\newcommand{\tuplee}[1]{\langle #1 \rangle}
\newcommand{\ALT}{~\mid~}
\newcommand{\lamb}[2]{\lambda #1.\,#2}
\newcommand{\spc}[0]{\quad}
\newcommand{\U}[0]{\cup}
\newcommand{\X}[0]{\times}
\newcommand{\xn}[0]{\cap}
\newcommand{\redsto}[0]{\longrightarrow}
\newcommand{\rredsto}[0]{\rightsquigarrow_r }
\newcommand{\evalsto}[0]{\longrightarrow }
\newcommand{\env}[0]{\Gamma}
\newcommand{\finmaparrow}[0]{\overset { Fin }{ \longrightarrow } }
\newcommand{\envwf}[1]{\vdash\,#1}
\newcommand{\tywf}[2]{#1\,\vdash\,#2}
\newcommand{\hastyp}[3]{#1\,\vdash\,#2\,:\,#3}
\newcommand{\hassort}[3]{#1\,\vdash\,#2\,::\,#3}
\newcommand{\subtyp}[3]{#1\,\vdash\,#2 <: #3}
\newcommand{\generalizes}[3]{#1\,\vdash\,#2 \succ #3}
\newcommand{\msfolsig}[0]{{\Lambda \equiv}}
%\newcommand{\msentails}[2]{#1\,\models_L\,#2}
\newcommand{\msentails}[2]{#1\,\models\,#2}
\newcommand{\thesemof}[1]{ \llbracket #1 \rrbracket}
\newcommand{\defeq}[0]{ \triangleq }
\newcommand{\tymap}[0]{\mathcal{F}}
\newcommand{\etaeq}[0]{\eta_{eq}}
\newcommand{\semof}[1]{\thesemof{#1}}
\newcommand{\mssemof}[1]{\thesemof{#1}}
\newcommand{\rwsto}[0]{=}
\newcommand{\rrwsto}[0]{\hookrightarrow}
\newcommand{\rwbindfwd}[0]{\gamma_{\Rightarrow}}
\newcommand{\rwbindbwd}[0]{\gamma_{\Leftarrow}}
\newcommand{\eqhlt}[1]{\underline{#1}}
\newcommand{\tfr}[0]{F_R}
\newcommand{\ground}[1]{\underline{#1}}
\newcommand{\semeq}[0]{\equiv}
\newcommand{\elabsto}[0]{\looparrowright }
\newcommand{\arr}[0]{$\rightarrow$}
\newcommand{\eef}[0]{$\Leftrightarrow$}
\newcommand{\imp}[0]{$\Rightarrow$}
\newcommand{\RNull}[0]{\emptyset}
\newcommand{\cty}[0]{ty}
\newcommand{\hsp}[0]{\hspace*{-0.1in}}
\newcommand{\hspa}[0]{\hspace*{-0.08in}}
% removable:
\newcommand{\subst}[2]{\lbrack #1/#2 \rbrack}
\newcommand{\rel}[1]{{R}_{\mathit{#1}}}
\newcommand{\absof}[1]{\left\| #1 \right\| }

@@

\lstset{ %
language=Java, % choose the language of the code
basicstyle=\scriptsize\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={relation, not, : , /\\}
}
\lstdefinestyle{hs} {
  language=Haskell,
  basicstyle=\scriptsize\ttfamily
}

@@

>>>>

\usepackage{courier}            % standard fixed width font
\usepackage[scaled]{helvet} % see www.ctan.org/get/macros/latex/required/psnfss/psnfss2e.pdf
\usepackage{listings}          % format code
\usepackage{enumitem}      % adjust spacing in enums
\usepackage{xspace}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{color}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{float}
\usepackage{verbatimbox}
\usepackage{breakurl}
\usepackage[breaklinks]{hyperref}

%Formatting Commands
%-------------------
\newcommand{\C}[1]{\code{#1}}
\renewcommand{\k}[1]{\mathtt{#1}} % Code font in math mode.
\newcommand{\code}[1]{\,{\tt #1}\,}
\newcommand{\conj}{\wedge}
\newcommand{\disj}{\vee}
\newcommand{\seqn}[1]{\overline {#1}}
\newcommand{\tuplee}[1]{\langle #1 \rangle}
\newcommand{\ALT}{~\mid~}
\newcommand{\lamb}[2]{\lambda #1.\,#2}
\newcommand{\spc}[0]{\quad}
\newcommand{\U}[0]{\cup}
\newcommand{\X}[0]{\times}
\newcommand{\xn}[0]{\cap}
\newcommand{\redsto}[0]{\longrightarrow}
\newcommand{\rredsto}[0]{\rightsquigarrow_r }
\newcommand{\evalsto}[0]{\longrightarrow }
\newcommand{\env}[0]{\Gamma}
\newcommand{\finmaparrow}[0]{\overset { Fin }{ \longrightarrow } }
\newcommand{\envwf}[1]{\vdash\,#1}
\newcommand{\tywf}[2]{#1\,\vdash\,#2}
\newcommand{\hastyp}[3]{#1\,\vdash\,#2\,:\,#3}
\newcommand{\hassort}[3]{#1\,\vdash\,#2\,::\,#3}
\newcommand{\subtyp}[3]{#1\,\vdash\,#2 <: #3}
\newcommand{\generalizes}[3]{#1\,\vdash\,#2 \succ #3}
\newcommand{\msfolsig}[0]{{\Lambda \equiv}}
%\newcommand{\msentails}[2]{#1\,\models_L\,#2}
\newcommand{\msentails}[2]{#1\,\models\,#2}
\newcommand{\thesemof}[1]{ \llbracket #1 \rrbracket}
\newcommand{\defeq}[0]{ \triangleq }
\newcommand{\tymap}[0]{\mathcal{F}}
\newcommand{\etaeq}[0]{\eta_{eq}}
\newcommand{\semof}[1]{\thesemof{#1}}
\newcommand{\mssemof}[1]{\thesemof{#1}}
\newcommand{\rwsto}[0]{=}
\newcommand{\rrwsto}[0]{\hookrightarrow}
\newcommand{\rwbindfwd}[0]{\gamma_{\Rightarrow}}
\newcommand{\rwbindbwd}[0]{\gamma_{\Leftarrow}}
\newcommand{\eqhlt}[1]{\underline{#1}}
\newcommand{\tfr}[0]{F_R}
\newcommand{\ground}[1]{\underline{#1}}
\newcommand{\semeq}[0]{\equiv}
\newcommand{\elabsto}[0]{\looparrowright }
\newcommand{\arr}[0]{$\rightarrow$}
\newcommand{\eef}[0]{$\Leftrightarrow$}
\newcommand{\imp}[0]{$\Rightarrow$}
\newcommand{\RNull}[0]{\emptyset}
\newcommand{\cty}[0]{ty}
\newcommand{\hsp}[0]{\hspace*{-0.1in}}
\newcommand{\hspa}[0]{\hspace*{-0.08in}}
% removable:
\newcommand{\subst}[2]{\lbrack #1/#2 \rbrack}
\newcommand{\rel}[1]{{R}_{\mathit{#1}}}
\newcommand{\absof}[1]{\left\| #1 \right\| }
\newcommand{\broom}{{\sc Broom}\xspace}
\newcommand{\name}{{\sc Broom}\xspace}
\newcommand{\naiad}{{\sc Naiad}\xspace}
\newcommand{\csolve}{{\sc Csolve}\xspace}
\newcommand{\GK}[1]{{\textcolor{red}{{GK: #1}}}}
\newcommand{\csharp}{C\#~}
\newcommand{\outlives}{\succeq}

@@

\newcommand{\lstjava}{\lstset{ %
language=Java, % choose the language of the code
basicstyle=\small\ttfamily,       % the size of the fonts that are used for the code
keywordstyle=\color{Bittersweet},
% numbers=left,                   % where to put the line-numbers
numberstyle=\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it is 1 each line will be numbered
numbersep=5pt,                  % how far the line-numbers are from the code
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
% frame=single,                   % adds a frame around the code
tabsize=2,                      % sets default tabsize to 2 spaces
captionpos=b,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
commentstyle=\itshape\color{MidnightBlue},
%escapeinside={\%*}{*)},         % if you want to add a comment within your code
morekeywords={letregion, openalloc, open, withroot, not, : , /\\}
}}
\lstnewenvironment{codejava}
    { % \centering
			\lstjava
      \lstset{}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

\lstnewenvironment{numcodejava}
    { % \centering
			\lstjava
      \lstset{numbers=left,firstnumber=1}%
      \csname lst@setfirstlabel\endcsname}
    { %\centering
      \csname lst@savefirstlabel\endcsname}

@@

\input{overview}

### 2015-09-12 paper/introduction.tex
Sec 2 full content.

\begin{itemize}
  \item We present \name, a Java-like typed object-oriented language
  that eschews garbage collection in favour of programmer-managed
  memory regions . \name extends its core language, which includes
  \emph{lambdas} (higher-order functions) and \emph{generics}
  (parametric polymorphism), with constructs to create, manage and
  destroy static and dynamic memory regions. Dynamic regions are
  first-class values in \name; they manifest as objects of type
  \C{Region}, and are treated on par with other objects. \name is
  particularly suited for implementing distributed dataflow operators
  (e.g: \c{SELECT} and \c{JOIN}), whose memory behavior exhibits
  certain characteristics that admits efficient and \emph{mostly} safe
  region-based memory management. Experiments demonstrate that
  \naiad~\cite{naiad} dataflow operators using programmer-managed
  memory regions outperform their GC counterparts by a margin of upto
  59\%.

@@

The rest of the paper is organized as follows. The next section
presents an informal overview of \name and its region type system.
\S~\ref{sec:type-system} formalizes the type system and its safety
guarantees. The type inference algorithm is described in
\S~\ref{sec:type-inference}. \S~\ref{sec:csolve} focuses on \csolve,
the constraint solving algorithm, and its correctness guarantees.
Implementations details and practical extensions to the type system
are described in \S~\ref{sec:implementation}. \S~\ref{sec:evaluation}
presents experimental evaluation and case studies.
\S~\ref{sec:related} discusses the related work, and
\S~\ref{sec:conclusion} concludes.

>>>>

\begin{itemize}
  \item We present \name, a \csharp-like typed object-oriented language
  that eschews garbage collection in favour of programmer-managed
  memory regions . \name extends its core language, which includes
  \emph{lambdas} (higher-order functions) and \emph{generics}
  (parametric polymorphism), with constructs to create, manage and
  destroy static and dynamic memory regions. Dynamic regions are
  first-class values in \name; they manifest as objects of type
  \C{Region}, and are treated on par with other objects. \name is
  particularly suited for implementing distributed dataflow operators
  (e.g: \c{SELECT} and \c{JOIN}), whose memory behavior exhibits
  certain characteristics that admits efficient and \emph{mostly} safe
  region-based memory management. Experiments demonstrate that
  \naiad~\cite{naiad} dataflow operators using programmer-managed
  memory regions outperform their GC counterparts by a margin of upto
  59\%.

@@

The rest of the paper is organized as follows. The next section
presents an informal overview of \name, and motivates the need for a
region type system.  \S~\ref{sec:type-system} formalizes the type
system and its safety guarantees. The type inference algorithm is
described in \S~\ref{sec:type-inference}. \S~\ref{sec:csolve} focuses
on \csolve, the constraint solving algorithm, and its correctness
guarantees.  Implementations details and practical extensions to the
type system are described in \S~\ref{sec:implementation}.
\S~\ref{sec:evaluation} presents experimental evaluation and case
studies.  \S~\ref{sec:related} discusses the related work, and
\S~\ref{sec:conclusion} concludes.
